The Mysterious Lives Of Speckles. I. Residual atmospheric speckle lifetimes behind AO-fed coronagraphs

### 2021-03-17 specklelives.tex
updated

\documentclass[11pt,preprint]{aastex}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{mathrsfs}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

@@

\usepackage{indentfirst}
\usepackage{lscape}
\usepackage{afterpage}
\usepackage{rotating}
\let\captionbox\undefined
%\usepackage{caption}

%\captionsetup{width=6.3in,font=footnotesize}
\usepackage{wrapfig}
\usepackage{multicol}
\usepackage{textcomp}
\newcommand{\textapprox}{\raisebox{0.5ex}{\texttildelow}}

\usepackage{multirow}

\usepackage{xcolor}
\usepackage{ulem}

\usepackage{soul}

\setlength{\textwidth}{6.5in}
\setlength{\hoffset}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}

\setlength{\textheight}{9.in}
\setlength{\voffset}{0in}
\setlength{\topmargin}{0in}
\setlength{\headsep}{0.25in}
\setlength{\headheight}{0in}

@@

\title{The Mysterious Lives Of Speckles. I. Residual atmospheric speckle lifetimes behind AO-fed coronagraphs}
\author{Jared Males}

@@

Characterization of extrasolar planets with resolved imaging is limited by the halo of starlight scattered onto the image plane by optical aberrations.  In addition to the photon noise caused by this halo, the scattered light is coherent and produces ``speckles'', which can be thought of as copies of the instrument point spread function (PSF).  This adds structured noise, with both spatial and temporal correlations, which is almost always the limiting noise source in a coronagraphic observation.    Both in space and on the ground, imperfections within the telescope and instrument optics cause speckles.   On the Earth's surface, the aberrations include those imposed by turbulence in the Earth's atmosphere.

The index of refraction of air depends on temperature, and turbulence causes the air temperature to stochastically vary across the telescope aperture.  As the stellar wavefront traverses these regions of different index of refraction, parts of the wavefront travel different optical path lengths.  Due mainly to wind, the atmospheric temperature structure is constantly changing.  The technique of adaptive optics (AO) is used to correct the resulting time-variable optical aberrations in real time.  Even with so-called ``extreme'' AO systems (ExAO, Guyon), this correction is not perfect.  The AO-corrected wavefront, though significantly improved, still causes a halo of speckles significantly brighter than the instrumental speckles in a modern high performance coronagraph system using wavefront control.

In order to analyze the capabilities of ground-based high-contrast imaging, we must understand the spatial and temporal evolution of the residual atmospheric speckles behind an ExAO coronagraph.  Speckle formation in the spatial-domain, as structures within the coronagraphic PSF, is well understood (xxxxx).  In the time-domain, we can imagine two ideal regimes for these structures.  In the first regime, the speckles are perfectly static.  Were this case, it would be a simple matter of measuring the PSF structure and then subtracting it from all images.  While this would still leave the photon noise from the scattered light, we could perfectly remove the spatially correlated noise.  In the second imaginary regime, the speckles are a perfect temporally white noise source: every image taken has a different speckle structure.  Similar to photon noise, it would now be a simple matter of taking enough images to average the speckles to a smooth halo.

Here we are concerned with the actual temporal behavior of atmospheric speckles, which vary on short time-scales making standard PSF-subtraction essentially impossible, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.  The statistical speckle lifetime, that is the time-scale which controls the averaging of speckle noise, is the quantity that sets the fundamental sensitivity limit of ground-based high-contrast imaging.

@@

More recent experimental studies have used on-sky data with modern ExAO systems.  Using focal plane data from SPHERE, \citet{2016SPIE.9909E..4ZM} found evidence for speckle decorrelation on several time-scales, but were unable to identify an atmospheric component given the various instrument-induced time-scales and the relatively low cadence used. In a study using SCExAO at much higher frame-rates, \citet{2018PASP..130j4502G} analyzed the short time-scale evolution of the intensity using the variance of difference images (essentially a form of autocorrelation).  Here short time-scale (a few ms) speckle evolution was attributed to the atmosphere, and the authors argued that the AO system modulated the speckles on longer time-scales.  \citet{2017JATIS...3b5001S} present a study of speckle lifetimes at the LBT, where they found short-lived speckles attributed to the atmosphere and show some evidence for a difference in lifetime inside and outside the AO control radius.

@@

\clearpage
\section{Discussion}
For some time, long-lived quasi-static speckles have been recognized as the limiting noise source in contemporary high contrast imaging observations \citep[e.g.][]{2007ApJ...654..633H}.  Given their origin within the instrument, and comparatively long lifetimes \citep{2012A&A...541A.136M}, such speckles should be supressable using wavefront sensing and control strategies.  Recent progress in applying ``dark hole'' contrast optimization with focal plane WFS \citep{2020A&A...638A.117P} and various low-order coronagraphic WFS\&C strategites \citep[e.g.][]{2017PASP..129i5002S} show great promise in significantly reducing the impact of internal speckles.  Assuming active quasi-static speckle control becomes routine, the fundamental sensitivity limit in ground-based high-contrast imaging is set by the Earth's atmosphere.  This limit is set by two components: uncorrelated photon noise and the spatially and temporally correlated speckle noise.  The temporal correlations of atmospheric speckles are, compared to quasi-static speckles, relatively short. This precludes conventional post-processing, but the lifetimes are long enough to significantly increase exposure times compared to the photon-noise.

>>>>

\documentclass[10pt,preprint]{aastex631}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{mathrsfs}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

@@

%\usepackage{indentfirst}
%\usepackage{lscape}
\usepackage{afterpage}
%\usepackage{rotating}
\let\captionbox\undefined
%\usepackage{caption}

%\captionsetup{width=6.3in,font=footnotesize}
%\usepackage{wrapfig}
%\usepackage{multicol}
%\usepackage{textcomp}
\newcommand{\textapprox}{\raisebox{0.5ex}{\texttildelow}}

%\usepackage{multirow}

%\usepackage{xcolor}
%\usepackage{ulem}

%\usepackage{soul}

%\setlength{\textwidth}{6.5in}
%\setlength{\hoffset}{0in}
%\setlength{\oddsidemargin}{0in}
%\setlength{\evensidemargin}{0in}

%\setlength{\textheight}{9.in}
%\setlength{\voffset}{0in}
%\setlength{\topmargin}{0in}
%\setlength{\headsep}{0.25in}
%\setlength{\headheight}{0in}

@@

\title{The Mysterious Lives Of Speckles. I. Residual atmospheric speckle lifetimes behind AO-fed coronagraphs}

\author{Jared R. Males}
\affiliation{Steward Observatory, University of Arizona, Tucson, 933 N Cherry Ave, Tucson, AZ 85721, USA}

\author{Michael P. Fitzgerald}
\affiliation{Department of Physics \& Astronomy, University of California, Los Angeles, CA 90095, USA}

\author{Ruslan Belikov}
\affiliation{NASA-Ames Research Center, Moffett Blvd., Moffett Field, CA, USA}

\author{Olivier Guyon}
\affiliation{Astrobiology Center, National Institutes of Natural Sciences, 2-21-1 Osawa, Mitaka, Tokyo, JAPAN}
\affiliation{Steward Observatory, University of Arizona, Tucson, 933 N Cherry Ave, Tucson, AZ 85721, USA}
\affiliation{College of Optical Science, University of Arizona, 1630 E University Blvd, Tucson, AZ 85719, USA}
\affiliation{National Astronomical Observatory of Japan, Subaru Telescope, National Institutes of Natural Sciences, Hilo, HI 96720, USA}

@@

Characterization of extrasolar planets with resolved imaging is limited by the halo of starlight scattered onto the image plane by optical aberrations.  In addition to the photon noise caused by this halo, the scattered light is coherent and produces ``speckles'', which can be thought of as copies of the instrument point spread function (PSF) \citep{1995PASP..107..386M}.  This adds structured noise, with both spatial and temporal correlations, which is the limiting noise source in a coronagraphic observation.    Both in space and on the ground, imperfections within the telescope and instrument optics cause speckles.   For telescopes on the Earth's surface, the aberrations include those imposed by turbulence in the Earth's atmosphere.

The index of refraction of air depends on temperature, and turbulence causes the air temperature to stochastically vary across the telescope aperture.  As the stellar wavefront traverses these regions of different index of refraction, parts of the wavefront travel different optical path lengths.  Due mainly to wind, the atmospheric temperature structure is constantly changing.  The technique of adaptive optics (AO) is used to correct the resulting time-variable optical aberrations in real time.  Even with so-called ``extreme'' AO systems \citep[ExAO, ][]{2018ARAA..56..315G} this correction is not perfect.  The AO-corrected wavefront, though significantly improved, still causes a halo of speckles significantly brighter than the instrumental speckles in a modern high performance coronagraph system using wavefront control.

In order to analyze the capabilities of ground-based high-contrast imaging, we must understand the spatial and temporal evolution of the residual atmospheric speckles behind an ExAO coronagraph.  Speckle formation in the spatial-domain, as structures within the coronagraphic PSF, is well understood \citep{2001ApJ...558L..71B,2003ApJ...596..702P,2007ApJ...669..642S}.  In the time-domain, we can imagine two ideal regimes for these structures.  In the first regime, the speckles are perfectly static.  Were this case, it would be a simple matter of measuring the PSF structure and then subtracting it from all images.  While this would still leave the photon noise from the scattered light, we could perfectly remove the spatially correlated noise.  In the second imaginary regime, the speckles are a perfect temporally white noise source: every image taken has a different speckle structure.  Similar to photon noise, it would now be a simple matter of taking enough images to average the speckles to a smooth halo.

In reality, the temporal behavior of atmospheric speckles falls between these regimes. These speckles vary on short time-scales making standard PSF-subtraction essentially impossible, but exhibit non-negligible temporal correlations which cause them to average much more slowly than photon noise.  The statistical speckle lifetime, that is the time-scale which controls the averaging of speckle noise, is the quantity that sets the fundamental sensitivity limit of ground-based high-contrast imaging.

@@

More recent experimental studies have used on-sky data with modern ExAO systems.  Using focal plane data from SPHERE, \citet{2016SPIE.9909E..4ZM} found evidence for speckle decorrelation on several time-scales, but were unable to identify an atmospheric component given the various instrument-induced time-scales and the relatively low cadence used. In a study using SCExAO at much higher frame-rates, \citet{2018PASP..130j4502G} analyzed the evolution of the intensity using the variance of difference images (essentially a form of autocorrelation).  Here short time-scale (a few ms) speckle evolution was attributed to the atmosphere, and the authors argued that the AO system modulated the speckles on longer time-scales.  \citet{2017JATIS...3b5001S} present a study of speckle lifetimes at the LBT, where they found short-lived speckles attributed to the atmosphere and show some evidence for a difference in lifetime inside and outside the AO control radius.

@@

\clearpage
\section{Discussion}
For some time, long-lived quasi-static speckles have been recognized as the limiting noise source in contemporary high contrast imaging observations \citep[e.g.][]{2007ApJ...654..633H}.  Given their origin within the instrument, and comparatively long lifetimes \citep{2012A&A...541A.136M}, such speckles should be supressable using wavefront sensing and control (WFS\&C) strategies.  Recent progress in applying ``dark hole'' contrast optimization with focal plane WFS \citep{2020A&A...638A.117P} and various focal-plane \citep{2010A&A...509A..31G} and low-order coronagraphic WFS\&C strategites \citep[e.g.][]{2017PASP..129i5002S} show great promise in significantly reducing the impact of internal speckles.  Assuming active quasi-static speckle control becomes routine, the fundamental sensitivity limit in ground-based high-contrast imaging is set by the Earth's atmosphere.  This limit is set by two components: uncorrelated photon noise and the spatially and temporally correlated speckle noise.  The temporal correlations of atmospheric speckles are, compared to quasi-static speckles, relatively short. This precludes conventional post-processing, but the lifetimes are long enough to significantly increase exposure times compared to the photon-noise.
